---
title: "solve_lin_lst_sqrs_for_ground_truthing"
author: "Blake Harlow"
date: "6/24/2021"
last edited: "6/30/2021"
output: pdf_document
---

```{r setup, include = FALSE}
gc()
rm(list=ls())
require("knitr")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

getwd()
```

```{r, import_libraries, warning = "hide", message = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readxl)
library(rgl)
library(seewave)  # has rms() function
library(pracma)   # has rref() function

source("code/createModelMatrix.R")
source("code/makeCleanData.R")
source("code/setOrigin.R")
```

Open the excel file as a tibble using readxl::read_excel()

```{r, read_data}
excel_path <- "raw_data/2021_06_23_ground_truthing.xlsx"
data <- read_excel(excel_path) %>%
  mutate(rn = row_number()) # This will enable me to track each sighting
```

Now, I am going to do a few trig checks to make sure that the data makes sense.

I will add horizontal_side_calc and vertical_side_calc columns that are found using hypotenuse and angle. Due to rounding differences, these are not exactly equal to horizontal_side and vertical_side, respectfully. However, I will filter out rows where one of them differ by more than 2%.


```{r, check_data}
data_check <- data %>%
  mutate(horizontal_side_calc = hypotenuse * cos(angle * pi / 180),
         vertical_side_calc = hypotenuse * sin(angle * pi / 180)) %>%
  filter((vertical_side_calc - vertical_side) / vertical_side > 0.02 | 
           (horizontal_side_calc - horizontal_side)/horizontal_side > 0.02) %>%
  select(page, laser_position, laser_height, target_position, target_height, 
         angle, hypotenuse, horizontal_side, 
         horizontal_side_calc, vertical_side, vertical_side_calc) %>%
  arrange((vertical_side_calc - vertical_side) / vertical_side) %>%
  select(horizontal_side, horizontal_side_calc, 
         vertical_side, vertical_side_calc)

# Now, look at this resulting tibble and see if there appear to be any blunders
data_check
```

There does not appear to be any major blunders in the data apparent from this check.

Next, I will construct the model matrix and the reponse vector (A and b in the equation Ax=b):

```{r, Construct_A_and_b}
# This is the order that I will assign for the laser and target indeces.
levels <- c("a20","a46","a53","a57","a61","a75","a76","a78","a87","a98",
            "t1","t2","t3","t4","t5","t6","t7","t8","t9","t10","t11","t12",
            "t13","t14","t15","t16",
            "p1","p2")

# First filter out aeropoint to aeropoint sightings and also add delta
# columns using sourced function:
clean_data <- makeCleanData(data)

# Now construct A using sourced function
A <- createModelMatrix(clean_data, levels) %>% setOrigin()
cnames <- colnames(A)

# Now construct b vectors:
b_x <- clean_data$delta_x %>% setOrigin()
b_y <- clean_data$delta_y %>% setOrigin()
b_z <- clean_data$delta_z %>% setOrigin()
```
Now that the A matrix and b vectors are constructed, we can do the fit:

```{r, fit_lm}
x_fit <- lm(b_x ~ A - 1, qr = TRUE)
y_fit <- lm(b_y ~ A - 1)
z_fit <- lm(b_z ~ A - 1)

cat("Residual Analysis of Original Data:\n\n",
    "The max residuals are:\n \t x:", max(x_fit$residuals),
    "\n \t y:", max(y_fit$residuals),
    "\n \t z:", max(z_fit$residuals),
    "\n\nThe rms of the residuals are:\n \t x:", rms(x_fit$residuals),
    "\n \t y:", rms(y_fit$residuals),
    "\n \t z:", rms(z_fit$residuals)
    )
```

Ok, why is there a giant x residual? I found the data measurement for this sighting that caused this residual (it was the sighting with clean_data index of 50). It was the sighting from t5 to t15 and had a recorded bearing of 92. This is clearly wrong. The sighting just after from t5 to t16 had a bearing of 307. Perhaps the 92 was supposed to be 292?

For now, I will remove that sighting from the data and re-do the least squares:



```{r, fit_lm2}
# First remove the problematic sighting (index 50)
data2 <- data %>% rows_delete(tibble(rn = 50))
clean_data2 <- makeCleanData(data2)

# Make a new A matrix and b vectors
A2 <- createModelMatrix(clean_data2, levels) %>% setOrigin()
cnames2 <- colnames(A2)
b_x2 <- clean_data2$delta_x %>% setOrigin()
b_y2 <- clean_data2$delta_y %>% setOrigin()
b_z2 <- clean_data2$delta_z %>% setOrigin()

# And now re-do the fit
x_fit2 <- lm(b_x2 ~ A2 - 1)
y_fit2 <- lm(b_y2 ~ A2 - 1)
z_fit2 <- lm(b_z2 ~ A2 - 1)


cat("Residual Analysis without the t5 to t15 sighting:\n\n",
    "The max residuals are:\n \t x:", max(x_fit2$residuals),
    "\n \t y:", max(y_fit2$residuals),
    "\n \t z:", max(z_fit2$residuals),
    "\n\nThe rms of the residuals are:\n \t x:", rms(x_fit2$residuals),
    "\n \t y:", rms(y_fit2$residuals),
    "\n \t z:", rms(z_fit2$residuals)
    )
```

I did check the sighting that provided the next highest x residual. It is the sighting from t14 to a75. I think that the bearing (297.5) is somewhat uncertain since it is greater than the bearing from t14 to a61 (291.5). From my map, I think the former bearing should have been less than the latter. However, I do not think my roughly drawn map is enough to warrant discarding this sighting.


What if I had changed the 92 degree bearing from t5 to t15 to 292 degrees rather than simply removing this point? Let's try:

```{r}
# First edit the row

data3 <- data %>% rows_update(tibble(rn = 50, bearing = 292.0)) %>%
  rows_delete(tibble(rn = 18)) %>%
  rows_delete(tibble(rn = 20)) %>%
  rows_delete(tibble(rn = 21)) %>%
  rows_delete(tibble(rn = 22)) %>%
  rows_delete(tibble(rn = 73))
clean_data3 <- makeCleanData(data3)

# make a new A matrix and b vectors
A3 <- createModelMatrix(clean_data3, levels) %>% setOrigin()
cnames3 <- colnames(A3)
b_x3 <- clean_data3$delta_x %>% setOrigin()
b_y3 <- clean_data3$delta_y %>% setOrigin()
b_z3 <- clean_data3$delta_z %>% setOrigin()

# And now re-do the fit
x_fit3 <- lm(b_x3 ~ A3 - 1)
y_fit3 <- lm(b_y3 ~ A3 - 1)
z_fit3 <- lm(b_z3 ~ A3 - 1)


cat("Residual Analysis with the edited t5 to t15 sighting:\n\n",
    "The max residuals are:\n \t x:", max(x_fit3$residuals),
    "\n \t y:", max(y_fit3$residuals),
    "\n \t z:", max(z_fit3$residuals),
    "\n\nThe rms of the residuals are:\n \t x:", rms(x_fit3$residuals),
    "\n \t y:", rms(y_fit3$residuals),
    "\n \t z:", rms(z_fit3$residuals)
    )

clean_data3 <- clean_data3 %>%
  mutate(x_residuals = x_fit3$residuals %>% head(-1),
         y_residuals = y_fit3$residuals %>% head(-1),
         z_residuals = z_fit3$residuals %>% head(-1),
         residuals = (x_residuals^2 + y_residuals^2 + z_residuals^2)^(1/2),
         x_start = x_fit3$coefficients[paste("A3", laser_position, sep = "")],
         y_start = y_fit3$coefficients[paste("A3", laser_position, sep = "")],
         z_start = z_fit3$coefficients[paste("A3", laser_position, sep = "")],
         x_end = x_start + delta_x,
         y_end = y_start + delta_y,
         z_end = z_start + delta_z) %>%
  filter(residuals > 2) # filtering out the sightings with greatest residuals

```


What if we only used the trap and points (discard sightings involving any aeropoints)? Let's try. Note that I am using data3, where the t5 to t15 bearing was adjusted to 292 degrees.

```{r}
# First remove aeropoint data:
data4 <- data3 %>% filter(substring(target_position,1,1) != "a")
clean_data4 <- makeCleanData(data4)

# make a new A matrix and b vectors
A4 <- createModelMatrix(clean_data4, levels) %>% setOrigin()
cnames4 <- colnames(A4)
b_x4 <- clean_data4$delta_x %>% setOrigin()
b_y4 <- clean_data4$delta_y %>% setOrigin()
b_z4 <- clean_data4$delta_z %>% setOrigin()

# And now re-do the fit
x_fit4 <- lm(b_x4 ~ A4 - 1)
y_fit4 <- lm(b_y4 ~ A4 - 1)
z_fit4 <- lm(b_z4 ~ A4 - 1)


cat("Residual Analysis without any aeropoints:\n\n",
    "The max residuals are:\n \t x:", max(x_fit4$residuals),
    "\n \t y:", max(y_fit4$residuals),
    "\n \t z:", max(z_fit4$residuals),
    "\n\nThe rms of the residuals are:\n \t x:", rms(x_fit4$residuals),
    "\n \t y:", rms(y_fit4$residuals),
    "\n \t z:", rms(z_fit4$residuals)
    )

```

Now, using the fit where I only changed the eroneous 92 bearing to 292, I will add a column for residuals to clean_data


Let's make a few plots: 

* A 3D plot showing both the 3rd fit as well as the 4th fit (without aeropoints). The points from each fit are distinguished with different colors. Note that this fit will not be visible in the knit output.
* A 2D plot showing:
  + The x-y location of all points from 3rd fit (the fit with the eroneous bearing adjusted from 92 to 292). The black text gives the label of each of these points.
  + Line segments revealing the sightings that have a total residual greater than 5 meters. The red text labels reveal the target of that sighting.

```{r, plot}
points3 <- tibble(position = cnames3,
                 x = x_fit3$coefficients, 
                 y = y_fit3$coefficients, 
                 z = z_fit3$coefficients,
                 is_aero = 1)

points4 <- tibble(position = cnames4,
                 x = x_fit4$coefficients, 
                 y = y_fit4$coefficients, 
                 z = z_fit4$coefficients,
                 is_aero = 2)

all_points <- rbind(points3, points4)

# Now try plotting. Note that these 3D plot are not visible in knit output
rgl.open()
rgl.bg(color = "white")
rgl.points(all_points$x, all_points$y, all_points$z, color = all_points$is_aero, size = 5)
bbox3d(color = "gray", alpha = 0.5)

# This plot should be visible in knit output
ggplot() +
  geom_segment(data = clean_data3,
              aes(x = x_start, y = y_start, xend = x_end, yend = y_end, 
                  color = residuals)) +
  geom_text(data = clean_data3, 
            aes(x_end, y_end, label = target_position), hjust=1, vjust=-1, 
            size = 3, color = "red") +
  geom_point(data = points3, 
             aes(x, y)) +
  geom_text(data = points3, 
            aes(x, y, label = position), hjust=-0.5, vjust=0, size = 3)
```
Notes:

t1 to a75 appears to a blunder.
bearings from t2 to t4, t5, t6, and p1 appear to be blunders


```{r}

points3

write.table(points3 %>% select(-is_aero), "./clean_data/initial_results.csv")

```



```{r, test_lm_using_numa_lab5, echo = FALSE, message = FALSE, results = "hide"}
# Here I am testing the lm() function with data that I have from NumA.
# I confirmed that these results are correct by looking back at lab 5.
# Because it agrees, the lm() function is doing exactly what I expect.

A_check = rbind(c(1,0),c(1,1),c(1,2),c(1,3),c(1,4),c(1,5))
b_check = c(1,2.7,5.8,6.6,7.5,9.9)

check1 <- lm(b_check ~ A_check - 1)

# The results of this agree with my results for my project. Nice.

A_check2 = rbind(c(1,0,0),c(1,1,1),c(1,2,4),c(1,3,9),c(1,4,16),c(1,5,25))
b_check2 = c(1,2.7,5.8,6.6,7.5,9.9)
check2 <- lm(b_check2 ~ A_check2 - 1)

message(cat(check1$coefficients, "should be [1.31904762, 1.70571429].\n"))
message(cat(check2$coefficients, "should be [ 1.00357143,  2.17892857, -0.09464286]."))
```